---
title: "CAnD3 IDEA 2: Disparities in Access to Health Care Services in Canada"
format: html
editor: visual
---

**Topic Choice:**

Disparities in Access to Health Care Services in Canada

**Research Question:**

What sociodemographic factors predict disparities in access to health care services in Canada?

## **Data Replicability & Workflow:**

-   Data Source: Odesi - [Canadian Community Health Survey, 2017-2018: Annual Component (odesi.ca)](https://odesi.ca/en/details?id=/odesi/doi__10-5683_SP3_EYLZ18.xml)

-   Data Set Download: GitHub 

-   Data Sharing: GitHub - <https://github.com/hi-polymer/cand3-idea2> 

-   Data Analysis: R Studio

## Data cleaning

### Loading data and libraries

```{r}
library(tidyverse)
library(haven)
```

Load data:

**Change address when replicating**

```{r}

cchs <- read_dta("/Users/thecommon/Downloads/cchs-82M0013-E-2017-2018-Annual-component_F1.dta")
```

### **Independent Variables:** 

-   Sex (DHH_SEX) 

-   Marital Status (DHHGMS) 

-   Age (DHHGAGE) 

-   Education (EHG2DVR3) 

-   Household Income (INCG015) 

-   Immigrant Status (SDCDVIMM) 

-   Visible Minority Status (SDCDGCGT)

-   Main activity last week (MACG005)

-   Occupational status (LBFDGOCG)

-   Survey weights (WTS_M)

```{r}
# variables for the demographic 
cchs <- cchs %>% mutate(sex = as_factor(DHH_SEX),
                marital_status = as_factor(DHHGMS),
                agegrp = as_factor(DHHGAGE),
                edu = as_factor(EHG2DVR3),
                hh_income = as_factor(INCDGHH),
                immigration = as_factor(SDCDVIMM),
                vis_min = as_factor(SDCDGCGT),
                occupation = as_factor(LBFDGOCG),
                main_activity = as_factor(MACG005),
                weights = WTS_M)

cchs %>% count(age)
# need to clean age to grab the digits
cchs <- cchs %>% mutate(
  agegrp = str_extract_all(agegrp, "([[:digit:]]{2})"),
  agegrp = map_chr(agegrp, str_c, collapse = "-"),  
  agegrp = if_else(agegrp == "80", "80+", agegrp)
) 


```

### **Dependent Variables:**

Inclusion flag (DOACC)

1.  Difficulties Accessing Health Care Services - composite index derived from:

    -   Experienced difficulties getting specialist care - 12 mo (ACC_010)

    -   Experienced difficulties getting non-emergency surgery - 12 mo (ACC_025)

    -   Experienced difficulties getting non-emergency test - 12 mo (ACC_040)

    -   Experienced difficulty getting health information - 12 mo (ACC_060)

    -   Experienced difficulties getting routine care - 12 mo (ACC_105)

    -   Experienced difficulties getting immediate care - 12 mo (ACC_135)

2.  No Access to Regular Health Care Provider

    -   Has a regular health care provider (ACC_095)

    -   **Instead we use PHC_020 that is administered to entire sample**

    -   Has a regular health care provider (PHC_020)

3.  Unmet healthcare needs (potential other outcome)

    -   Unmet healthcare needs (DOUCN)

    -   self_perception (UCN_005)

    -   reasons (UCN_010A to K)

    -   type of care (UCN_015A to J)

```{r}

cchs <- cchs %>% mutate(
  acc_module = as_factor(DOACC),
  specialist = as_factor(ACC_005),
  specialist_diff = as_factor(ACC_010),
  surgery = as_factor(ACC_020),
  surgery_diff = as_factor(ACC_025),
  medtest = as_factor(ACC_035),
  medtest_diff = as_factor(ACC_040),
  healthinfo = as_factor(ACC_050),
  healthinfo_diff = as_factor(ACC_060),
  routinecare = as_factor(ACC_100),
  routinecare_diff = as_factor(ACC_105),
  immidiatecare = as_factor(ACC_130),
  immidiatecare_diff = as_factor(ACC_135),
  ACC_regulardoc = as_factor(ACC_095),
  regulardoc = as_factor(PHC_020)
)

# Checking where the module was administered
cchs %>% count(acc_module, as_factor(GEO_PRV))


```

This module was only administered to Nova Scotia and Nunavut.

#### Constructing the index variable:

```{r}
# constructing the variable for difficulity

cchs <- cchs %>% mutate(
  across(contains("_diff"), ~ if_else(.x == "Yes", 1, 0), .names = "{.col}_bin" 
  ),
  # increasing number of difficulities
  num_diff = rowSums(pick(ends_with("_diff_bin")),na.rm = TRUE),
  num_diff = if_else(acc_module == "No", NA, num_diff),
  # by any difficulities
  any_diff = if_else(num_diff>0, "Yes", "No")
  ) 

cchs %>% count(acc_module,any_diff)
```

Main dependent variable: any_diff

### Save file

**Change address when replicating**

```{# Change address when replicating}
{r}
cchs_sm <- cchs %>% select(GEO_PRV,sex:agegrp,starts_with("ACC"),
                           DOUCN,starts_with("UCN"))

rm(cchs)
save(cchs_sm,file = "/Users/thecommon/Documents/cchs_2017_18.Rda")

```
